---
title: "Misunderstanding Hegemony"
format: html
---

## Setup

```{r packages + settings}
library(tidyverse)
library(socsci)
library(geomtextpath)
library(ggrepel)
library(coolorrr)
library(here)
set_palette(
  qualitative = c(
    "steelblue",
    "navy",
    "red3",
    "orange3"
  ),
  binary = c("orange3", "navy"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "navy")
)
source(here::here("my_theme.R"))
theme_set(my_theme())
saveit <- function(num = 1, ht = 4, wd = 6) {
  ggsave(
    here(
      "048_hegemony",
      paste0("fig", num, ".png")
    ),
    height = ht,
    width = wd,
    dpi = 500
  )
}
```

## Data

I need to do my level best to replicate the security hierarchy measure proposed in Lake (2007). What I need:

1. US Troop deployments: Total troops deployed in a country divided by that country's population then normalized to a 0-1 scale.

2. US alliance dependence: 1 divided by the number of "independent alliances" --- alliances with other countries that are not also allied with the US. If the country isn't allied with the US at all, then its alliance dependence value is zero.

Let's get the data:

```{r}
## alliances

## 1. start with a dyadic alliance-year dataset
library(peacesciencer)
create_dyadyears(
  subset_years = 1950:2014
) |>
  add_atop_alliance() -> ad

## 2. Aggregate to get the alliance dependence measure
ad |>
  
  # make indicator for a non-ally of the US
  group_by(ccode2, year) |>
  mutate(
    nonus2 = max(
      1 - atop_defense[ccode1 == 2],
      na.rm = T
    )
  ) |> 
  
  # count up alliance independence per country
  group_by(ccode1, year) |>
  mutate(
    us1 = max(
      atop_defense[ccode2 == 2],
      na.rm = T
    )
  ) |> 
  filter(
    ccode1 != 2 &
      ccode2 != 2
  ) |>
  summarize(
    us1 = max(us1),
    sumind = sum(
      nonus2 * atop_defense
    ),
    ai = case_when(
      us1 == 0 ~ 0,
      us1 == 1 & sumind > 0 ~ 1 / sumind,
      us1 == 1 & sumind == 0 ~ 1
    )
  ) |>
  ungroup() |>
  transmute(
    ccode = ccode1,
    year, ai
  ) -> sad

## now troop data
library(troopdata)

td <- get_troopdata(
  endyear = 2014
)

td |>
  transmute(
    region,
    ccode, 
    year,
    troops = troops_ad
  ) -> td

cow_nmc |>
  transmute(
    ccode, year,
    pop = 1000 * tpop,
    milex
  ) |>
  right_join(
    td
  ) |>
  filter(ccode != 2) |>
  mutate(
    ti = (troops / pop) / max(troops / pop, na.rm = T)
  ) -> td

## merge together
fd <- left_join(
  sad, td
)

## now use to populate a dataset of conflict onsets
md <- read_csv(
  here("_data", "mie-1.0.csv")
)

md |>
  filter(hostlev >= 4, sidea1 == 1) |>
  distinct(ccode1, styear) |>
  transmute(
    ccode = ccode1,
    year = styear,
    mic_init = 1
  ) -> md

fd |>
  left_join(
    md
  ) |>
  mutate(
    mic_init = replace_na(mic_init, 0)
  ) -> fd
```

Whew! That took some extra brain power to make those variables work.

## Figures

Okay, what does the distribution of these indexes look like?

First I gotta make the combined index:

```{r}
fd |>
  mutate(
    secind = (ai + ti) / max(ai + ti, na.rm = T)
  ) -> fd
```


Okay, now I can plot:

```{r}
#| fig-height: 5
#| fig-width: 6

fd |>
  select(year, ai, ti, secind) |>
  pivot_longer(-year) |>
  mutate(
    name = frcode(
      name == "secind" ~ "Combined",
      name == "ai" ~ "Alliance",
      name == "ti" ~ "Troops"
    )
  ) |>
  ggplot() +
  aes(year, value, color = name) +
  stat_smooth(
    method = "gam",
    method.args = list(quasibinomial)
  ) +
  ggpal() +
  labs(
    title = str_wrap(
      "Measures of US security hierarchy from 1950 to 2014",
      width = 37
    ),
    subtitle = str_wrap(
      "Indexes based on Lake (2007)"
    ),
    x = NULL,
    y = "Mean Hierarchy",
    caption = logo,
    color = "Index"
  ) +
  theme(
    legend.title.position = "top",
    legend.title = element_text(
      hjust = .5
    )
  )
saveit(1, 5)
```

```{r}
#| fig-height: 5
#| fig-width: 6

ggplot(fd) +
  aes(secind, milex / 1e06) +
  stat_smooth(
    method = "glm",
    method.args = list(family = quasipoisson),
    level = .834,
    color = "navy"
  ) +
  scale_y_continuous(
    labels = ~ paste0("$", comma(.x), " bil")
  ) +
  labs(
    title = str_wrap(
      "Security subordination to the US predicts lower military spending",
      width = 37
    ),
    subtitle = str_wrap(
      "Predicted defense spending using a pooled psuedo-Poisson model. 83.5% CIs shown.",
      width = 55
    ),
    x = "Security Index",
    y = NULL,
    caption = logo
  )
saveit(2, 5)
```

```{r}
#| fig-height: 5
#| fig-width: 6

ggplot(fd) +
  aes(secind, mic_init) +
  stat_smooth(
    method = "glm",
    method.args = list(family = binomial),
    level = .834,
    color = "navy"
  ) +
  scale_y_continuous(
    labels = percent
  ) +
  labs(
    title = str_wrap(
      "Security subordination to the US predicts lower conflict propensity",
      width = 37
    ),
    subtitle = str_wrap(
      "Predicted chance of conflict initiation using a pooled logit model. 83.5% CIs shown.",
      width = 55
    ),
    x = "Security Index",
    y = NULL,
    caption = logo
  )
saveit(3, 5)
```


Notes:

* Consistent with Lake (2007), US hegemony predicts both lower military spending by subordinates and lower propensity to initiate international conflicts.
* But US hegemony is on the decline, and has been since the beginning of the Cold War. 
* This might be one explanation for the current increase in global conflict --- the less subordinate countries are to a security hegemon, the more they take their own security into their own hands, and the more conflict we see in the world as a result.
* Case in point is the current posture of America's NATO allies. They now doubt that the US is a reliable guarantor of their security, so they are ramping up military spending. 
* The point of being a hegemon, of bearing the security burden of other countries, is to ensure that the world remains a more peaceful place. The reality of a world where our allies "pay up" is one where the US has less dominance and international conflict is more common.