---
title: "WW1 vs WW2"
format: html
---

## Setup

```{r packages + settings}
library(tidyverse)
library(socsci)
library(geomtextpath)
library(coolorrr)
library(ggrepel)
library(here)
set_palette(
  qualitative = c(
    "orange3",
    "navy",
    "steelblue"
  ),
  binary = c("orange3", "navy"),
  sequential = c("white", "navy"),
  diverging = c("orange3", "white", "navy")
)
source(here::here("my_theme.R"))
theme_set(my_theme())
saveit <- function(num = 1, ht = 4, wd = 6) {
  ggsave(
    here::here(
      "086_world_wars",
      paste0("fig", num, ".png")
    ),
    height = ht,
    width = wd,
    dpi = 500
  )
}
```

## Data

```{r}
## good thing I made this for my students already:
source("https://raw.githubusercontent.com/milesdwilliams15/death-destruction-data/refs/heads/main/helpers/get_mie_data.R")
get_mie_data() -> dt

dt |>
  mutate(
    date2 = paste0(endyear, "-", endmon) |> ym()
  ) -> dt

dt |>
  filter(micnum %in% 257:257) |>
  mutate(
    diff = date2 - date
  ) |>
  count(diff) |>
  mutate(pct = n / sum(n)) 
```

```{r}
dt |>
  group_by(micnum) |>
  summarize(
    fatal = sum(fatalmin1 + fatalmin2)
  ) |>
  mutate(
    ww = ifelse(micnum %in% c(257, 258), 1, 0)
  ) |>
  group_by(ww) |>
  summarize(
    fatal = sum(fatal)
  ) |>
  ungroup() |>
  mutate(
    pct = fatal / sum(fatal)
  )
```


```{r}
## just the world wars
dt |>
  filter(micnum %in% 257:258) |>
  mutate(
    war = ifelse(micnum == 257, "WW1", "WW2")
  ) -> dt
```

## Figures

```{r}
dt |>
  count(war)
```


```{r}
#| fig-height: 4.5
#| fig-width: 7
dt |>
  group_by(war) |>
  count(date) |>
  ggplot() +
  aes(date, n) +
  geom_line(
    aes(color = "Monthly Total"),
    linewidth = .75
  ) +
  geom_smooth(
    aes(color = "Overall Average"),
    method = "lm",
    se = F,
    formula = y ~ 1
  ) +
  geom_smooth(
    aes(color = "Smoothed Trend"),
    method = "lm",
    formula = y ~ poly(x, 3),
    se = F,
    linetype = 2
  ) +
  facet_wrap(~ war, scales = "free_x") +
  labs(
    title = "WW1 saw more battles per month than WW2",
    subtitle = "Battles Initiated per Month (MIE dataset version 1.0)",
    x = NULL,
    y = NULL,
    caption = logo,
    color = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 500, by = 25)
  ) +
  ggpal()
saveit(1, 4.5, 7)
```

Total battles:

```{r}
dt |>
  count(war)
```


```{r}
#| fig-height: 4.5
#| fig-width: 7
dt |>
  group_by(war, date2) |>
  summarize(
    n = sum(fatalmax1 + fatalmax2)
  ) |>
  ggplot() +
  aes(date2, n) +
  geom_line(
    aes(color = "Monthly Total"),
    linewidth = .75
  ) +
  geom_smooth(
    aes(color = "Overall Average"),
    method = "lm",
    se = F,
    formula = y ~ 1
  ) +
  geom_smooth(
    aes(color = "Smoothed Trend"),
    method = "lm",
    formula = y ~ poly(x, 3),
    se = F,
    linetype = 2
  ) +
  facet_wrap(~ war, scales = "free_x") +
  labs(
    title = "Both wars saw about as many deaths per month\n(but WW2 witnessed bigger extremes)",
    subtitle = "Millions of Battle Deaths per Month (MIE dataset version 1.0)",
    x = NULL,
    y = NULL,
    caption = logo,
    color = NULL
  ) +
  scale_y_continuous(
    breaks = seq(0, 2e06, by = 1e05),
    labels = ~ scales::comma(.x / 1e06)
  ) +
  ggpal()
saveit(2, 4.5, 7)
```

Total deaths:

```{r}
dt |>
  group_by(war) |>
  summarize(min = sum(fatalmin1 + fatalmin2), max = sum(fatalmax1 + fatalmax2))
```



```{r}
#| fig-height: 4.5
#| fig-width: 7
## death per battle
source(here("_helpers", "my_mean_ci.R"))
dt |>
  group_by(war) |>
  summarize(
    my_mean_ci(fatalmax1 + fatalmax2)
  ) |> 
  ggplot() +
  aes(war, y, ymin = ymin, ymax = ymax) +
  geom_pointrange(color = "navy") +
  labs(
    title = "Deaths per battle were much higher in WW2",
    subtitle = "Average deaths per battle (MIE dataset version 1.0)",
    x = NULL,
    y = NULL,
    caption = logo
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  theme(panel.grid.major.x = element_blank())
saveit(3, 4.5, 7)
```

Some big takeaways:

1. WW1 saw a more pronounced upward trend in war battles and far more battles overall (thanks to trench warfare) than WW2.
2. WW2 saw a more pronounced upward trend in battle deaths and more deaths per battle overall (thanks to developments in military technology) than WW1.
3. More fighting doesn't always equate to more deaths. Technology and the style of warfare matter, too.

```{r}
325000 / ((2026 - (2022 + 2/12)) * 12)
```

```{r}
dt |>
  group_by(war, date) |>
  summarize(
    sum = sum(fatalmax1 + fatalmax2)
  ) |>
  group_by(war) |>
  summarize(
    mean = mean(sum)
  )
```

