---
title: "2024 presidential election: county-level exploration"
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.
:::

## Setup

```{r}
#| label: load-packages
library(tidyverse)
library(readxl)
library(usmap)
```

```{r}
#| label: load-data
election <- read_excel(
  "../_data/2024_Presidential_Election_Results/Presidential Election Results by US County 2024.xlsx"
)
```

## Data structure

The dataset contains county-level (and county-equivalent) results for the 2024 U.S. presidential election, along with the winning candidate in 2024, 2020, and 2016.

```{r}
#| label: glimpse-data
glimpse(election)
```

The data covers `{r} n_distinct(election$STATE_NAME)` states/territories across `{r} nrow(election)` rows.

```{r}
#| label: winner-counts
election |>
  count(Winner_2024)
```

Excluding rows with missing vote counts, the national two-party-plus totals are:

```{r}
#| label: vote-totals
election |>
  summarise(
    total_votes = sum(Votes_Tot, na.rm = TRUE),
    trump = sum(Votes_Trump, na.rm = TRUE),
    harris = sum(Votes_Harris, na.rm = TRUE),
    stein = sum(Votes_Stein, na.rm = TRUE)
  )
```

## Missing data

Vote totals, vote counts by candidate, and vote shares are each missing for exactly one county:

```{r}
#| label: missing-votes
election |>
  filter(is.na(Votes_Tot))
```

This is **Kalawao County, Hawaii** — the smallest county in the U.S. by population (a former Hansen's disease settlement on Molokai). Its missing values are plausibly a genuine data gap rather than an entry error, given how few residents it has.

Separately, `Winner_2020` and `Winner_2016` are missing for a larger set of counties:

```{r}
#| label: missing-historical-winners
election |>
  summarise(across(c(Winner_2024, Winner_2020, Winner_2016), ~ sum(is.na(.))))
```

```{r}
#| label: missing-historical-winners-by-state
election |>
  filter(is.na(Winner_2020) | is.na(Winner_2016)) |>
  count(STATE_NAME)
```

These missing historical winners are concentrated in Virginia (38 of 39 counties), with one additional case in Hawaii. This pattern looks systematic rather than random — it likely reflects how Virginia's independent cities are handled in the source data across election years — and is worth further investigation before using the historical winner columns.

## County-level margin map

We compute the two-party vote margin for each county as the Trump minus Harris vote share, using only Trump and Harris votes in the denominator (i.e., excluding Stein and other third-party votes):

```{r}
#| label: compute-margin
county_margin <- election |>
  filter(!is.na(Votes_Trump), !is.na(Votes_Harris)) |>
  mutate(
    two_party_total = Votes_Trump + Votes_Harris,
    margin = (Votes_Trump - Votes_Harris) / two_party_total,
    fips = FIPS
  )
```

```{r}
#| label: fig-county-margin-map
#| fig-height: 6
#| fig-width: 8
plot_usmap(
  regions = "counties",
  data = county_margin,
  values = "margin",
  color = NA
) +
  scale_fill_gradient2(
    low = "#2166AC", mid = "white", high = "#B2182B",
    midpoint = 0,
    labels = scales::label_percent(),
    name = "Margin (Trump \u2212 Harris)"
  ) +
  labs(title = "2024 presidential election: county-level two-party vote margin") +
  theme(legend.position = "right")
```

Two limitations to note when reading this map:

- **Alaska is not shown.** The source data contains no Alaska rows at all — Alaska reports results by state house district rather than by county/borough, so there is nothing to join to the map.
- **A small gray sliver** (visible in the Northeast on this projection) corresponds to Kalawao County, HI, which has no vote data as noted above.
