---
title: "Untitled"
format: html
---

## Setup

```{r packages + settings}
library(tidyverse)
library(socsci)
library(geomtextpath)
library(coolorrr)
library(ggrepel)
library(here)
library(countrycode)
set_palette(
  qualitative = c(
    "orange3",
    "steelblue",
    "navy",
    "red3",
    "gray"
  ),
  binary = c("orange3", "navy"),
  sequential = c("white", "navy"),
  diverging = c("orange3", "white", "navy")
)
source(here::here("my_theme.R"))
theme_set(my_theme())
saveit <- function(num = 1, ht = 4, wd = 6) {
  ggsave(
    here::here(
      "088_medal_counts",
      paste0("fig", num, ".png")
    ),
    height = ht,
    width = wd,
    dpi = 500
  )
}
```



## Data

Source: https://www.kaggle.com/datasets/heesoo37/120-years-of-olympic-history-athletes-and-results?resource=download

```{r}
read_csv(
  here("_data", "athlete_events.csv")
) -> dt

dt |>
  group_by(Team, Year, Season) |>
  summarize(
    n_ath = n_distinct(ID)
  ) -> n_aths

dt |>
  group_by(Team, Year, Season, Event) |>
  summarize(
    gold = any(Medal == "Gold") + 0,
    silver = any(Medal == "Silver") + 0,
    bronze = any(Medal == "Bronze") + 0,
    .groups = "drop"
  ) |>
  group_by(Team, Year, Season) |>
  summarize(
    across(gold:bronze, ~ sum(.x, na.rm = T)),
    total = gold + silver + bronze,
    .groups = "drop"
  ) -> n_medals

left_join(
  n_aths,
  n_medals
) -> n_sum

n_sum |>
  distinct(
    Team
  ) -> teams

teams |>
  mutate(
    ccode = countrycode(Team, "country.name", "cown")
  ) -> teams

n_sum |>
  mutate(
    ccode = countrycode(Team, "country.name", "cown")
  ) |>
  drop_na(ccode) |>
  select(ccode, everything()) |>
  janitor::clean_names() -> n_sum
```

```{r}
## okay so multiple teams per country. I can fix that.
n_sum |>
  group_by(ccode, year) |>
  filter(n_ath == max(n_ath)) |>
  ungroup() -> n_sum

## bring in peacescience data
library(peacesciencer)
create_stateyears() |>
  add_sim_gdp_pop() |>
  add_democracy() -> ps_dt

n_sum |>
  left_join(
    ps_dt
  ) -> final_dt
```

Let's make this sucker publically available:

```{r}
write_csv(
  dt,
  here("_data", "olympics_data.csv")
)
```


## Figures

```{r}
#| fig-height: 5
#| fig-width: 7

ggplot(dt) +
  aes(year, total, color = season) +
  geom_path(
    aes(group = ccode),
    linewidth = .1
  ) +
  geom_path(
    data = . %>% filter(ccode == 2),
    linewidth = 1
  ) +
  facet_wrap(~ season) +
  ggpal("binary") +
  theme(
    legend.position = ""
  ) +
  labs(
    title = str_wrap(
      "The US regularly dominates in the Olympics"
    ),
    subtitle = str_wrap(
      "Total medal counts by country (U.S. in bold), 1896-2016"
    ),
    x = NULL,
    y = NULL,
    caption = logo
  )
saveit(1, 5, 7)
```


```{r}
#| fig-height: 5
#| fig-width: 7

ggplot(dt) +
  aes(year, n_ath, color = season) +
  geom_path(
    aes(group = ccode),
    linewidth = .1
  ) +
  geom_path(
    data = . %>% filter(ccode == 2),
    linewidth = 1
  ) +
  facet_wrap(~ season) +
  ggpal("binary") +
  theme(
    legend.position = ""
  ) +
  labs(
    title = str_wrap(
      "The US sends large Olympic delegations"
    ),
    subtitle = str_wrap(
      "Total athletes by country (U.S. in bold), 1896-2016"
    ),
    x = NULL,
    y = NULL,
    caption = logo
  )
saveit(2, 5, 7)
```

```{r}
#| fig-height: 5
#| fig-width: 7

ggplot(dt) +
  aes(n_ath, total) +
  geom_point(color = "gray") +
  geom_smooth(
    method = "gam",
    se = F,
    color = "navy"
  ) +
  labs(
    title = "A larger delegation = more medals",
    x = "Delegation",
    y = "Medals",
    caption = logo
  )
saveit(3, 5, 7)
```


```{r}
#| fig-height: 5
#| fig-width: 7

library(mgcv) 
gam(
  total ~ log(n_ath) +
    log(pwtpop) +
    log(pwtrgdp/pwtpop) +
    log(v2x_polyarchy) +
    season +
    poly(year, 3) +
    s(ccode, bs = "re"),
  data = dt,
  family = quasipoisson
) -> fit

summary(fit)
```
```{r}
#| fig-height: 5
#| fig-width: 7

fit |>
  lmtest::coeftest() |>
  broom::tidy() |>
  mutate(
    lower = estimate - 1.96 * std.error,
    upper = estimate + 1.96 * std.error
  ) |>
  slice(2:5) |>
  ggplot() +
  aes(
    estimate, term, xmin = lower, xmax = upper
  ) +
  geom_pointrange() +
  geom_vline(xintercept = 0, lty = 2) +
  labs(
    title = "Which countries get an edge in medals?",
    x = "PPML estimates\n(season, time, and random recipient effects not shown)",
    subtitle = "% change in medals per a % change in predictors",
    y = NULL,
    caption = logo
  ) +
  scale_y_discrete(
    labels = c(
      "Delegation (ln)",
      "Population (ln)",
      "GDP per capita (ln)",
      "Democracy"
    )
  ) +
  scale_x_continuous(
    labels = ~ paste0(.x, "%")
  )
saveit(4, 5, 7)
```


```{r}
gam(
  n_ath ~
    log(pwtpop) +
    log(pwtrgdp / pwtpop) +
    log(v2x_polyarchy) +
    season +
    poly(year, 3) +
    s(ccode, bs = "re"),
  data = dt,
  family = quasipoisson
) -> fit2

summary(fit2)
```

```{r}
#| fig-height: 5
#| fig-width: 7

fit2 |>
  lmtest::coeftest() |>
  broom::tidy() |>
  mutate(
    lower = estimate - 1.96 * std.error,
    upper = estimate + 1.96 * std.error
  ) |>
  slice(2:4) |>
  ggplot() +
  aes(
    estimate, term, xmin = lower, xmax = upper
  ) +
  geom_pointrange() +
  geom_vline(xintercept = 0, lty = 2) +
  labs(
    title = "Which countries send bigger delegations?",
    x = "PPML estimates\n(season, time, and random recipient effects not shown)",
    subtitle = "% change in delegates per a % change in predictors",
    y = NULL,
    caption = logo
  ) +
  scale_y_discrete(
    labels = c(
      "Population (ln)",
      "GDP per capita (ln)",
      "Democracy"
    )
  ) +
  scale_x_continuous(
    labels = ~ paste0(.x, "%")
  )
saveit(5, 5, 7)
```




Notes:

* Why does no one have a free and easy to access country-year dataset of Olympic medal counts??
* US does disproportionately well
* US sends a lot of athletes (so easiest explanation for performance is a numbers game)
* Other factors make a marginal difference, too: population (unexpectedly negative), size of the economy (positive), and democracy (unexpectedly negative).
* Democracy is odd, but it also predicts an ability to send a larger delegation so...

